### 2024-04-24 paper/paper.qmd
paper update

---
title: "[WIP] Church Bells are Not Ringing: Decrease in the Opposite-Sex and Same-Sex Couples Opting for Religous Marriage Ceremonies in England and Wales"
subtitle: "Analysis and Comparative Study of Marriage Ceremony Type (Civil and Religious) for Opposite-Sex (1970 to 2020) and Same-Sex (2016 to 2020) Couples "
author: 
  - Francesca Ye
thanks: "Code and data are available at: https://github.com/Francesca-Ye/marriage_ceremonies_religion"
date: today
date-format: long
abstract: "This paper investigates the annual number of marriage conducted in England and Wales based on ceremony type recorded the by the Office of National Statistics. Based on the data set, these trends between 1970 to 2020 were analyzed for opposite-sex couples and between 2016 to 2020 for same-sex couples in tandem with comparative study on couple type preferences. Analysis found that civil ceremonies have overtake civil cermonies as the predominant ceremony type for both couple types. The results of this paper indicate that religion is less of an inetgral facet of life in England and Wales as indicated by religion becoming less synonymous with marriage."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---
```{r}
#| message: false
#| echo: false
#| warning: false
#### Work space set-up ####

## Load packages
library(tinytex)
library(tidyverse)
library(ggplot2)
library(knitr)
library(arrow)
library(tidyr)

## Read in cleaned data 
opposite_sex <- read_parquet("~/marriage_ceremonies_religion/data/analysis_data/00-cleaned_opposite_sex_data.parquet")

same_sex <- read_parquet("~/marriage_ceremonies_religion/data/analysis_data/01-cleaned_same_sex_data.parquet")
```

# Introduction {#sec-intro}
In the United Kingdom, there has been increasing research and attention towards the decline of the importance of religion in individual life. The Policy Institute at King’s College London has collected data which corroborates this trend. In 1981, 57% of Britons identified as religious while that number has decreased to 32% in 2022. Similarly, 75% of Britons believed in God when surveyed in 1981 and that number has significantly fallen to 49% in 2022 [@ukreligion]. However, these trends are not unique to the United Kingdom, religiosity across most Western countries has declined over the past couple of decades. The number of people in the United States that identified as religious fell from 82% in 1982 to 58% in 2017 [@ukreligion].

Despite the overall decline in religiosity, there are also generational differences in religious belief; as noted in the United Kingdom where only 48% of Baby Boomers expressed a belief in God in comparison to Gen Z at 37% [@ukreligion]. These generational differences and societal changes can be examined through data on marriage pertaining to the annual number of civil and religious ceremonies conducted. Although there is a degree of separation between civil and religious marriage, the practice has held heavy religious significance historically. Therefore, the types of marriage ceremonies people are opting for can serve as a representation of the overall religiosity of a country depending on the number of citizens that continue to ascribe religious significance to the legal act of marriage [@marriagedeclineuk].

This paper takes data from the *Office for National Statistics in the United Kingdom* from the 2020 edition of the "Marriages in England and Wales" data set to examine if the presence of religion in legal marriage has declined over time in tandem with answering if the result is difference between same and opposite sex couples. The estimand of interest is the presence of religion in consolidating a legal marriage. To examine whether or not religion in marriage has declined, graphical analysis was conducted to visualize the annual data on marriage ceremony type between 1970 to 2020 for opposite-sex couples and 2016 to 2020 for same-sex couples. Summary statistics were also calculated to get a greater idea of the numerical significance of religion in this space.

This paper finds that religion in the context of marriage has declined over the years in addition to same-sex couples more strongly preferring civil ceremonies in comparison to opposite-sex couples. More specifically, religious ceremonies appears to be rapidly declining in popularity amongst opposite-sex couples over the past five decades whilst civil ceremonies appear to be comparatively more stable in annual totals. This could imply that religiosity has decreased in England and Wales as couples are choosing to not involve religion in a significant life event. Interestingly, same-sex couples have strongly preferred civil ceremonies since the legalization of same-sex marriage in 2014 which makes a difference between opposite-sex couples. The latter has seen an active decline whilst the former has maintained an aversion to religious marriage ceremonies. Although both result in a fewer religious ceremonies, this indicates that the move away from religion could be motivated by different factors such as homophobia within religion for same-sex couples.

[@sec-data] of this report aims to discuss the characteristics of the data set examined in addition to the methods of cleaning applied to get workable data for this report. This report’s [@sec-results] highlights trends discovered through graphical analysis in tandem with statistical analysis. [@sec-discussion] notes possible analytical limitations and measurement errors in the data sets in addition to overall insights of the report’s analysis. Specific discussion will surround the validity of using marriage as an indicator of religiosity in further detail. [@sec-future] aims to explore further research that could be conducted in this field based on the findings presented in this report.

# Data {#sec-data}
## Data Description
The data sets for same and opposite-sex couples were derived from the *Office for National Statistics in the United Kingdom* with the 2020 edition of the “Marriages in England and Wales” workbook as raw data. This data is collected at the time a marriage is registered as part of civil registration which is a legal requirement in completing a legally recognized marriage in England and Wales. Annual marriage statistics are available to the public approximately two years after the end of the reference year [@statsguide]. Notably, same-sex couples that are currently in a civil partnership can have it converted to a marriage but these numbers are not counted towards the marriage statistics of same-sex couples in the workbook. The data collected and used in this paper is used to measure the number of couples that opt for civil or religious marriage ceremonies on a yearly basis [@statsguide].

By measuring marriage ceremony trends, it can help measure and provide insight onto greater social attitudes towards religion. Based on aforementioned literature about the perceived decline of religiosity in the United Kingdom, marriage ceremony data was chosen as a potential form of measuring this phenomena due to the historical link between marriage and religion [@marriagedeclineuk]. Since marriage is a significant life event, the presence and absence of religion in ceremonies can corroborate the decline of religiosity in individual life. These marriage ceremony entries in the "Marriages in England and Wales" workbook are a reflection of each couple's relationship with religion and whether or not it is strong enough to opt for a religious ceremony. Although the data set only reflects ceremony types, those entries were antedated by personal consideration of a couple's proximity to religion.

The data of this paper focuses on the ceremony statistics for both same and opposite-sex couples. For same-sex couples, the data set focuses on 2014-2020. The variables in this data set are `all_marriages` which represents the total amount of same-sex marriages conducted in a specific year, `all_civil` which represents all civil ceremonies conducted, `approved_civil` which represents all civil ceremonies conducted at premises with pre-approval, and `all_religious` which represents all religious ceremonies conducted. For same-sex couples, the number of male and female couples that marry in a year is also recorded. For opposite-sex couples, the data set focuses on 1970-2020 since there was a larger set of data available for analysis in comparison to same-sex couple data. The variables in this data set are `all_marriages` which represents the total amount of same-sex marriages conducted in a specific year, `all_civil` which represents all civil ceremonies conducted, `approved_civil` which represents all civil ceremonies conducted at premises with pre-approval, and `all_religious` which represents all religious ceremonies conducted. Religious ceremonies are further separated with annual ceremonies conducted in the Church of England, Roman Catholic churches, other Christian religions, and all other religions also recorded.

## Data Cleaning
The data set was imported into RStudio where the open-source statistical programming language, R [@citeR], was used for cleaning and analysis.  Data was simulated with the functionalities of the `tibble` [@tibble] package. The raw data set was extracted and saved with the functionalities of the `readr` [@readr], `tidyverse` [@tidyverse], and `readxl` [@readxl] packages. Data cleaning was aided by the use of the `tidyverse` [@tidyverse], `janitor` [@janitor], `arrow` [@arrow], and `naniar` [@naniar] packages. To create tables and graphs for data visualization of the datasets, `ggplot2` [@ggplot2], `tidyr` [@tidyr], and `knitr` [@knitr] packages were used in tandem with `tinytex` [@tinytex] to create the paper format.

Generally, the data from the *National Office of Statistics* was taken without any modifications to most of the selected columns. The notable exception to this would be in the opposite-sex couple dataset where the `approved_civil` column had values of `[z]` and `[x]` in the `other_religions` column when no data was available for a specific year. Since this paper aimed to broadly look at a summary of religion in marriage ceremonies, these values were converted to 0s to continue calculating summary statistics for individual years. All of the values in both data sets were also converted from the character class to the numeric class which is a divergence from the original workbook produced by the *National Office of Statistics*. This decision was made to aid in calculating and graphing the data as there were no values that could not be converted to numeric aside from the two cases mentioned above. Samples of the cleaned datasets produced are seen in @tbl-opposite and @tbl-same.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-opposite
#| tbl-cap: Sample of Cleaned Opposite-Sex Marriage Ceremony Data

# Code referenced from: https://github.com/InessaDeAngelis/Toronto_Elections/blob/main/outputs/paper/toronto_elections_paper.qmd

head (opposite_sex, 5) |>
  kable(
    col.names = c("Year","Total Marriages","All Civil", "Approved Civil", "All Religious", "Church of England", "Roman Catholic", "Other Christian Denominations", "Other Religions"),
  booktabs = TRUE
  )
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-same
#| tbl-cap: Cleaned Same-Sex Marriage Ceremony Data

# Code referenced from: https://github.com/InessaDeAngelis/Toronto_Elections/blob/main/outputs/paper/toronto_elections_paper.qmd
head (same_sex) |>
  kable(
    col.names = c("Year","Total Marriages","Male", "Female", "All Civil", "Approved Civil", "All Religious"),
  booktabs = TRUE
  )
```

## Opposite-Sex Civil versus Religious Ceremonies
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-oppositecivilvsreligious
#| fig-pos: "H"
#| fig-cap: Annual Amount of Civil (Red) and Religious (Blue) Ceremonies Between 1970 to 2020 in England and Wales for Opposite-Sex Couples

ggplot(opposite_sex, aes(x = year)) + 
  geom_line(aes(y = all_civil), color = "darkred") + 
  geom_line(aes(y = all_religious), color="steelblue")  +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1970, 2020, 5)) +
  scale_y_continuous(breaks = seq(0, 300000, 20000))+
  labs(x = "Year", y = "Number of Ceremonies Conducted")
```

As seen in @fig-oppositecivilvsreligious, the number of annual religious marriage ceremonies appears to have a clear downward trend from 1970 to 2020 which is contrasted by annual civil ceremonies that appear to have maintained somewhat more consistent levels throughout the 50 years. Interestingly, civil ceremonies were more popular than religious ceremonies between 1976 and 1980. However, from 1992, civil ceremonies were once again more popular and have consistently stayed ahead of religious ceremonies. After 1992, the gap between the amount of civil and religious ceremonies also appears to be exponentially widening as exemplified by 2015 where 17,6406 civil ceremonies were conducted in comparison to 6,2614 religious ceremonies. When solely viewing the annual total of religious and civil ceremonies conducted annually, it paints a clear picture of the decrease in religious marriage ceremonies. For both ceremony types, the decrease between 2019 and 2020 could be attributed to COVID-19-related lockdowns.

The average number of civil ceremonies conducted annually was 168,306.6 whilst the average for religious ceremonies was 130,945.3. Additionally, the medians were 168,897 and 127,522 respectively. The standard deviation for civil ceremonies was 17,601.7 and 59,287.21 for religious ceremonies. Since the standard deviation of religious ceremonies is significantly higher, this corroborates @fig-oppositecivilvsreligious in supporting that the annual totals have had large variances whereas civil ceremonies have maintained a comparatively stable annual total.

## Same-Sex Civil versus Religious Ceremonies
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-samecivilvsreligious
#| fig-pos: "H"
#| fig-cap: Annual Amount of Civil (Red) and Religious (Blue) Ceremonies Between 2014 and 2020 in England and Wales for Same-Sex Couples

ggplot(same_sex, aes(x = year)) + 
  geom_line(aes(y = all_civil), color = "darkred") + 
  geom_line(aes(y = all_religious), color="steelblue")  +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2014, 2020, 1)) +
  scale_y_continuous(breaks = seq(0, 8000, 500))+
  labs(x = "Year", y = "Number of Ceremonies Conducted")
```

As shown in @fig-samecivilvsreligious, the number of civil ceremonies has consistently been the more popular choice for ceremony type amongst same-sex couples since the legalization of same-sex marriage. Despite the smaller number of same-sex marriages in general, civil ceremonies make up an overwhelming majority of all marriages conducted annually. Although civil ceremonies are consistently the most popular, there does appear to be some slight year-over-year decrease that might have some significance; however, the decrease between 2019 and 2020 could be attributed to COVID-19-related lockdowns. Looking over the religious ceremonies, they have appeared to remain somewhat consistent with the exception of 2014 where only 17 ceremonies were conducted. Despite this, @fig-samecivilvsreligious highlights how civil ceremonies are the overwhelmingly popular choice for same-sex couples when selecting a marriage ceremony type.

For civil ceremonies, the average amount of annual ceremonies conducted was 6036.6 with a median of 6862. The average amount of annual religious ceremonies conducted was 46.4 with a median of 48. For standard deviations, civil ceremonies were 1815.576 and 18.48783 for religious ceremonies. In this instance, the larger variation in annual totals for civil ceremonies could be a result of the larger volume of ceremonies conducted. Since religious ceremonies conducted annually have consistently been in the double digits, the low standard deviation makes sense when viewing @fig-samecivilvsreligious as the line appears to be trending as an almost straight line.

# Results {#sec-results}
Looking at the graphed data and summary statistics, there appears to be a clear decline in the number of opposite-sex couples opting for a religious ceremony across a fifty-year time frame between 1970 and 2020. For same-sex couples, civil ceremonies have consistently and overwhelmingly been the most popular ceremony type since the legalization of same-sex marriage in 2014.Looking further into the data sets, it is apparent that preliminary conclusions from the summary data in @sec-data hold true after further analysis.

## Opposite-Sex Religious Ceremony Type
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-oppositereligioustype
#| fig-pos: "H"
#| fig-cap: Annual Amount of Church of England (Red), Roman Catholic (Yellow), Other Christian Denominations (Green), and Other Relgion Ceremonies (Purple) Conducted Between 1970 and 2020 in England and Wales for Opposite-Sex Couples

ggplot(opposite_sex, aes(x = year)) + 
  geom_line(aes(y = coe), color = "darkred") + 
  geom_line(aes(y = rc), color="yellow")  +
  geom_line(aes(y = other_christian), color="darkgreen")  +
  geom_line(aes(y = other_religion), color="darkviolet")  +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1970, 2020, 5)) +
  scale_y_continuous(breaks = seq(0, 175000, 15000))+
  labs(x = "Year", y = "Number of Ceremonies Conducted")
```

While @fig-oppositecivilvsreligious shows that the annual total of religious ceremonies as a whole have decreased, the annual total of of religious ceremonies sub categorized by type reveal a similar story. @fig-oppositereligioustype highlights that all Christian denominations have seen a decrease in the annual total of ceremonies conducted. While all other religions appears to be an outlier in this instance, the number of non-Christian religious ceremonies conducted has consistently been the least conducted type of ceremony amongst opposite-sex couples. Additionally, since the collection of data on non-Christian religious ceremonies, the standard deviation is relatively small at 810.76 which further highlights the minimal impact non-Christian ceremonies have on the larger data set combined with their unpopularity. As the only category with growth for religious ceremonies, this could mean that religiosity is not growing immensely outside of religiosity.

It is also apparent that Christianity has decreased in popularity for marriage ceremonies. Comparatively, Roman Catholic and other Christian denominations have seen decreases over a fifty-year period to a significant extent. The standard deviation for Roman Catholic is 10,788.53 and 11,781.59 for other Christian denominations which is considerably varied and highlights the significant drop in marriage ceremonies of these sects as seen in graphical analysis. However, the most notable drop would be Church of England ceremonies which had a standard deviation of 38,034.88, the largest, and when factoring in the downward graphical trend, it is clear that the Church of England has experienced major decreases in popularity for marriage ceremonies. Resultantly, Christian religious ceremonies have experienced somewhat extreme decreases in popularity that could imply changing individual and social attitudes on religion.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-oppositereligioustypeten
#| fig-pos: "H"
#| fig-cap: Annual Amount of Church of England (Red), Roman Catholic (Yellow), Other Christian Denominations (Green), and Other Relgion Ceremonies (Purple) Conducted Between 2011 and 2020 in England and Wales for Opposite-Sex Couples

ten_year <- subset(opposite_sex, year >= 2011 & year <= 2020)
ggplot(ten_year, aes(x = year)) + 
  geom_line(aes(y = coe), color = "darkred") + 
  geom_line(aes(y = rc), color = "yellow")  +
  geom_line(aes(y = other_christian), color="darkgreen")  +
  geom_line(aes(y = other_religion), color="darkviolet")  +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2011, 2020, 1)) +
  scale_y_continuous(breaks = seq(0, 60000, 5000))+
  labs(x = "Year", y = "Number of Ceremonies Conducted")
```

By looking at the most recent decade, the decline in religious ceremonies still appears to be rapid. In @fig-oppositereligioustypeten, the peak of the decade for Church of England ceremonies is 2012 with 79,473; however, the lowest year excluding 2020 is 2019 with 39,897 ceremonies. There is almost a 40,000 ceremony drop between these two years which further indicate the decline of religious ceremonies and rapid pacing just over a ten-year period. As the most popular religion for marriage ceremonies by a considerable margin, the Church of England still appears to be experiencing significant drops in popularity.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-oppositedecades
#| tbl-cap: Total Amount of Marriage Ceremony Types Conducted by Decade (1971-2020)

ten_year1 <- subset(opposite_sex, year >= 1971 & year <= 1980)
ten_year2 <- subset(opposite_sex, year >= 1981 & year <= 1990)
ten_year3 <- subset(opposite_sex, year >= 1991 & year <= 2000)
ten_year4 <- subset(opposite_sex, year >= 2001 & year <= 2010)

opposite_decades <-
  data.frame(
  type = c("Civil", "Religious"),
  one = c(sum(ten_year1$all_civil), sum(ten_year1$all_religious)),
  two = c(sum(ten_year2$all_civil), sum(ten_year2$all_religious)),
  three = c(sum(ten_year3$all_civil), sum(ten_year3$all_religious)),
  four = c(sum(ten_year4$all_civil), sum(ten_year4$all_religious)),
  five = c(sum(ten_year$all_civil), sum(ten_year$all_religious)))

# Code referenced from: https://github.com/InessaDeAngelis/Toronto_Elections/blob/main/outputs/paper/toronto_elections_paper.qmd
head (opposite_decades) |>
  kable(
    col.names = c("Ceremony Type","1971-1980","1981-1990", "1991-2000", "2001-2010", "2011-2020"),
  booktabs = TRUE
  )
```

@tbl-oppositedecades shows the total number of religious and civil ceremonies across five decades between 1971 and 2020. In the 1970s, religious ceremonies are more popular by just under 200,000 more ceremonies conducted. While both ceremony types see a decrease in the 1980s, the gap between the two is only about 100,000 with religious ceremonies still leading. Interestingly, civil ceremonies appear to minimally fluctuate by 100,000 ceremonies over the next three decades whilst religious ceremonies actively decrease in their decade summary. Religious ceremonies conducted in the 1970s was 1,995,792 in comparison to 567,934 in the 2010s. This is contrasted by civil ceremonies which saw 1,823,284 and 1,674,720 in the very same decades. Considering the decade summaries, it is clear that civil ceremonies have not necessarily increased by taking away from couples that might have opted for religious ceremonies due to their relative stability in numbers over time. It can also be ascertained that religious ceremonies have experienced a severe decreased over the past five decades as the annual total of ceremonies being conducted presently is just over a quarter to the number of ceremonies in the 1970s.

## Same-Sex Couple Ceremony Totals
```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-samesexdemographics
#| fig-pos: "H"
#| fig-cap: Annual Amount of Marriage Ceremonies Conducted Between 2014 and 2020 in England and Wales for Opposite-Sex Couples by Sex

same_sex_demographics <-
  pivot_longer(same_sex, cols = c(male, female), names_to = "Sex", values_to = "Count")

ggplot(same_sex_demographics, aes(x = year, y = Count, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year",
       y = "Number of Married Couples") +
  scale_fill_manual(values = c("darkred", "steelblue"),
                    labels = c("Female", "Male")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 4000, 500))
```

As seen in @fig-samecivilvsreligious, civil ceremonies for same-sex couples have consistently been the premier ceremony choice for couples intending to marry. @fig-samesexdemographics visualizes the annual number of couples that marry by sex. As shown in @fig-samesexdemographics, female couples have consistently married more than male couples in any given year. However, the gap between these totals are not large with 2019 representing the largest gap of 996 marriages. It can be derived that sex does not play a significant factor in the popularity of civil ceremonies as females do not make up a large majority of all same-sex married couples. Therefore, the unpopularity of religious ceremonies could be due to other factors which is touched upon in @sec-discussion.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-samesextotals
#| tbl-cap: Total Amount of Same-Sex Marriage Ceremonies Conducted Between 2014 to 2020

same_sex_numbers <-
  data.frame(
    type = c("Civil", "Religious"),
    value = c(sum(same_sex$all_civil), sum(same_sex$all_religious))
  )

# Code referenced from: https://github.com/InessaDeAngelis/Toronto_Elections/blob/main/outputs/paper/toronto_elections_paper.qmd
head (same_sex_numbers) |>
  kable(
    col.names = c("Ceremony Type", "Total Number of Ceremonies Conducted"),
  booktabs = TRUE
  )

```

As same-sex marriage was only recently legalized, the six years of data was summed up by civil and religious to provide a greater insight on the preferences of same-sex couples. @tbl-samesextotals provides the total number of religious and civil ceremonies conducted for same-sex couples since 2014 and up to 2020. Civil ceremonies are clearly the more popular choice with just under 30,000 more ceremonies conducted. Of all same-sex marriage ceremonies conducted since the the inception of legalization, religious ceremonies only make up about 0.76% whilst civil ceremonies make up the rest. This differs with opposite-sex marriages as they have experienced a decline of religious ceremonies whereas religious ceremonies for same-sex marriages have been extremely unpopular from the outset of legalization.

# Discussion {#sec-discussion}
## Implication: Religiosity has Decreased in England and Wales
## Implication: Fewer Citizens Identify as Christian
## Implication: Same-Sex Couples Overwhelmingly Prefer Civil Ceremonies
## Limitations
## Future Research {#sec-future}

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References